 \stepcounter{lecture}
 \setcounter{lecture}{28}
 \sektion{Lecture 28}

 \begin{definition}
   A ring $R$ is called \emph{good} (called \emph{property $\ast$} in the notes) if for
   every non-unit $a\in R$, $\bigcap (a^i)=0$.
 \end{definition}
 ``good'' is a necessary condition for certain normality properties.
 \begin{proposition}
   $R$ is good if any of the following hold.
   \begin{enumerate}
     \item $R$ is a UFD.
     \item $R$ is a noetherian domain.
     \item $R$ is completely normal.
     \item $R[[x]]$ is completely normal.
   \end{enumerate}
 \end{proposition}
 \begin{proof}
   1 and 2 we've done already \anton{}. Assume $R$ is a bad domain, so $0\neq d=\bigcap
   (a^i)$ for some $a\in R$. Then $d(a^{-1})^i\in R$ for all $i$. Then $a^{-1}\not\in R$
   is almost integral over $R$, contradicting complete normality.

   Believe for the moment that there is an $f=x+b_2x^2+\cdots \in \ZZ[[x]]$ such that
   $f^2-f+x=0$. Now replace $x$ by $\frac{x}{a^2}$. Then we have
   $f(x/a^2)^2-f(x/a)+x/a^2=0$, so if we define $g:= a\cdot f(x/a^2)$, we have
   $g^2-ag+x=0$, so $g$ is integral over $A=R[[x]]$ (even over $R[x]$). Note that $dg\in
   A$ since $d$ ``clears'' the denominators, so $g\in Q{A}$. But $g\not\in A$, which we
   can see by expanding it (we get a $1/a$ coefficient).

   Finally, the $f$ above can be constructed by looking at the constraints on the $b_i$,
   or solve with the quadratic equation ... you get catalan numbers as coefficients.
 \end{proof}

 picture
 \[\xymatrix{
  &R\ UFD \ar[d]\\
  R\ cn\ar[r] \ar@/^/[rr] & R\ g & R\ n \ar@{.>}[l]|{\times}\\
  R[[x]]\ cn \ar@{<->}[u] \ar[rr] &  & R[[x]]\ n \ar[lu] \ar[u]\\
   & R\ ndom \ar[uu]
 }\]

 \begin{example}[a normal $R$ with $A=R[[x]]$ not normal]
   Take $R_0=\QQ[x,y]\subseteq \QQ[x,y/x]\subseteq \QQ[x,y/x^2]\subseteq \cdots$, so
   $R_i=\QQ[x,y/x^i]$. Consider $R= \bigcup R_i$. Each $R_i\cong \QQ[x,t]$ is normal,
   from which it is easy to prove that $R$ is normal, but $x^i|y$ for all $i$, and $x$ is
   not a unit. So $R$ is bad.
 \end{example}
 For another example, refine the $D+(x)$ idea. Take $D$ a normal domain, with
 $D\subsetneq K=Q(D)$. Define $R= D+(x)\subseteq K[x]$.
 \begin{proposition}
   $R$ is normal and has complete integral closure $R^\dag = K[x]$ and is bad.
 \end{proposition}
 \begin{proof}
   Suppose $g\in K(x)=Q(R)$ is integral over $R$, so $g^n+f_1g^{n-1}+\cdots + f_n=0$,
   with $f_i\in R$. Evaluating at 0, we see that $g(0)$ is integral over $D$, so $g(0)\in
   D$, so $g\in R$. Thus, $R$ is normal.

   Take a non-unit $a\in D$, then $x/a^i\in R$ for all $i$, so $x\in \bigcap (a^i)$, so
   $R$ is bad.
 \end{proof}
